---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(scipen=999)
```

# rowcolr

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of `rowcolr` is to extract structured data from Excel files by identifying row and column labels using regex patterns or predefined identifiers. The package is designed to facilitate easy extraction of data by locating values at the intersection of row and column labels, making data wrangling and cleaning more efficient.

`rowcolr` leverages the excellent `tidyxl` package to read data from Excel files, ensuring robust and reliable extraction of cell contents. The package is particularly useful for dealing with semi-structured data in Excel files, which can be difficult to process using existing packages.

## Origin

The development of `rowcolr` was driven by a specific use case: existing Excel files used as input forms by organizations in the arts and culture sector to provide information to the Dutch Ministry of Education, Culture, and Science (OCW). To minimize disruptions to the existing workflow, the forms were kept as similar as possible to their original format while still allowing structured extraction of all relevant data from the Excel files.

## Installation

You can install the development version of `rowcolr` from GitHub with:

``` r
devtools::install_github("coeneisma/rowcolr")
```

## Usage

The primary function of `rowcolr` is `extract_values()`, which allows you to extract structured data from Excel files by identifying row and column labels. The function works by drawing a horizontal line from a detected row label and a vertical line from a detected column label. The value located at their intersection is extracted and stored.

The package can be loaded with:

```{r}
library(rowcolr)
```

### Example: Extracting Data from an Excel File

Values can be extracted from an Excel file using the default row and column regex patterns `.*_row$` and `.*_col$`:

```{r}
# Extract values from an example Excel file
dataset <- extract_values(rowcolr_example("example.xlsx"))

head(dataset |> 
       dplyr::select(-c(filename, sheet, row, col)))
```

For more details on using the package, refer to the vignette: `vignette("rowcolr")`. 
