---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(scipen=999)
```

# rowcolr

<!-- badges: start -->

<!-- badges: end -->

The goal of `rowcolr` is to extract structured data from Excel files by identifying row and column labels using regex patterns or predefined identifiers. The package is designed to support easy extraction of data based on matching row and column labels, facilitating data wrangling and cleaning tasks from spreadsheet data.

## Origin

The development of `rowcolr` was driven by a specific use case: existing Excel files used as input forms by organizations in the arts and culture sector to provide information to the Dutch Ministry of Education, Culture, and Science (OCW). To minimize disruptions to the existing workflow, the forms were kept as similar as possible to their original format, while still allowing structured extraction of all relevant data from the Excel files.

## Installation

You can install the development version of `rowcolr` from GitHub with:

``` r
devtools::install_github("coeneisma/rowcolr")
```

## Usage

The primary function of `rowcolr` is `extract_values()`, which allows you to extract structured data from Excel files based on row and column label patterns.

The package can be loaded with:

```{r}
library(rowcolr)
```


### Example: Extracting Data from an Excel File

Values can be extract from an Excel file using default row/column regex patterns `.*_row$` and .*_col$`:

```{r}
# 
dataset <- extract_values(rowcolr_example("example.xlsx"))

head(dataset |> 
       dplyr::select(-c(filename, sheet, row, col)))
```

More on the use of the package can be found in the vignette: `vignette("rowcolr")`.
