---
title: "rowcolr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rowcolr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Usage

The only function of `rowcolr` is `extract_values()`, which extracts structured data from Excel files by identifying row and column labels. The function works by drawing a horizontal line from a detected row label and a vertical line from a detected column label. The value at their intersection is extracted and stored.

The function provides **flexibility** by allowing both **explicit identifiers** and **regex patterns**:

-   If explicit identifiers (`row_identifiers` and `col_identifiers`) are provided, they take priority over regex.

-   If regex patterns (`row_pattern`, `col_pattern`) are provided, they act as a fallback when explicit identifiers are not exhaustive.

The package can be loaded with:

```{r setup}
library(rowcolr)
```

### Extracting Data from an Excel File using specific row and column identifiers

If you have predefined row or column labels, you can use them directly by providing character vectors to the `col_identifiers` and `row_identifiers` arguments:

```{r}
row_identifiers_example <- c("int_fixed_assets_row", "tang_fixed_assets_row",
                             "fin_fixed_assets_row", "tot_fixed_assets_row")
col_identifiers_example <- "2025_col"

dataset <- extract_values(rowcolr_example("example.xlsx"), 
                          row_identifiers = row_identifiers_example,
                          col_identifiers = col_identifiers_example)

head(dataset |> dplyr::select(-c(filename, sheet, row, col)))

```

### Extracting Data from an Excel File using regex patterns

Values can be extracted from an Excel file using regex patterns:

```{r}
# Extract values using regex patterns
dataset <- extract_values(rowcolr_example("example.xlsx"), 
                          row_pattern = ".*_row$", 
                          col_pattern = ".*_col$")

head(dataset |> dplyr::select(-c(filename, sheet, row, col)))

```

### Extracting Data from an Excel File using both identifiers and regex patterns 

You can combine the use of predefined row or column labels and regex patterns:

```{r}
dataset <- extract_values(rowcolr_example("example.xlsx"), 
                          row_identifiers = row_identifiers_example,
                          col_identifiers = col_identifiers_example,
                          row_pattern = ".*_row$", 
                          col_pattern = ".*_col$")
```



### No regex pattern or identifiers

If **no identifiers or regex are provided**, `extract_values()` will return an **error**:

``` r
dataset <- extract_values(rowcolr_example("example.xlsx"))

#> Error: No identifiers or regex patterns provided. Please specify 
#> `row_identifiers`, `col_identifiers`, `row_pattern`, or `col_pattern`.
```

### Controlling Description Cleaning

By default, row and column labels are cleaned by removing row and column patterns before combining them into the description. To disable this, use `clean_description = FALSE`:

```{r}
dataset <- extract_values(rowcolr_example("example.xlsx"), 
                          row_pattern = ".*_row$", 
                          col_pattern = ".*_col$",
                          clean_description = FALSE)

head(dataset |> 
       dplyr::select(c(row_label, col_label, description)))
```

## Conclusion

The `rowcolr` package is a powerful tool for extracting structured data from Excel files based on customizable row and column patterns. It is designed to be flexible, allowing for both regex-based matching and predefined lists of identifiers.
