---
title: "rowcolr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rowcolr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rowcolr)
```

### Example: Extracting Data from an Excel File

``` r
library(rowcolr)

# Extract values from an Excel file using default row/column regex patterns
dataset <- extract_values("path/to/your/excel_file.xlsx")
head(dataset)
```

### Example: Using Specific Row and Column Identifiers

If you have predefined row or column labels, you can use them directly by providing a character vector.

``` r
# Extract values using predefined row and column identifiers
dataset <- extract_values("path/to/your/excel_file.xlsx",
                           row_identifiers = c("Total Assets", "Net Profit"),
                           col_identifiers = c("2023", "2024"))
head(dataset)
```

### Example: Cleaning Descriptions by Removing Suffix Patterns

You can also clean the row and column labels before combining them into the description.

``` r
# Extract values with cleaned descriptions
dataset <- extract_values("path/to/your/excel_file.xlsx", clean_description = TRUE)
head(dataset)
```

## Conclusion

The `rowcolr` package is a powerful tool for extracting structured data from Excel files based on customizable row and column patterns. It is designed to be flexible, allowing for both regex-based matching and predefined lists of identifiers.
